---
title: "INFLUENZA A H7N9 IN CHINA, 2013"
format: dashboard
theme: lux

---

```{r}
#load packages 

pacman::p_load(tidyverse, outbreaks,
               bslib, bsicons, shiny, 
               rnaturalearth, plotly, 
               countrycode, htmltools, 
               reactable, sf
               )


```

```{r}
# read dataset
flu_china <- outbreaks::fluH7N9_china_2013

# Calculate Total Number of Cases

total_cases <- flu_china %>% 
  nrow()

#calculate total deaths 
total_deaths <- flu_china %>% 
  filter(outcome == "Death") %>% 
  nrow()

#calculate number of hospitalisation

total_hospitalisation <- 
  flu_china %>% 
  filter(!is.na(date_of_hospitalisation)) %>% 
  nrow()


```

```{r}
# create bar plot

cases_province_summary <- flu_china %>%
  group_by(province) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>% 
  mutate(tooltip_label = paste(province), 
                               sep = ": ")

cases_province <- ggplot(cases_province_summary, aes(y = reorder(province, count), x = count,
                                                     fill = count,
                                                     text = tooltip_label)) +
  geom_bar(stat = "identity", fill = "#3f71ab") +
  geom_text(aes(label = province),
            nudge_x = -10,
            color = "white"
            ) +
  labs(x = "Number of Cases", y = "Province") +
  theme(legend.position = "none")

plot_flu <- ggplotly(cases_province, tooltip = "text")

```

```{r}
province_summary <- fluH7N9_china_2013 %>% 
  group_by(province) %>% 
  summarise(total_cases = n())

country_shapes <- rnaturalearth::ne_states(country = "China", returnclass = "sf") 

# Get the administrative boundaries (provinces) for China

china <- ne_countries(country = "China", returnclass = "sf")
china_map <- ne_states(country = "China", returnclass = "sf")



# Check the keys to see if join will be successful
setdiff(province_summary$province, china_map$shapeName)
setdiff(china_map$shapeName, province_summary$province)
# Some work to be done to harmonize these

china_map <- china_map %>%
  # Clean up province names by removing unwanted terms
  mutate(province = str_trim(
    str_remove(shapeName, "Province|Municipality|Autonomous Region")
  ))

# Check if join will work fine
setdiff(province_summary$province, china_map$province)

# Fix Guangdong, which based on a quick Google is actually Guangzhou
china_map <- china_map %>% 
  mutate(province = if_else(province == "Guangzhou", "Guangdong", province)) 
  
# Check again
setdiff(province_summary$province, china_map$province)

# Now join
china_map_joined <- 
  left_join(china_map, province_summary, by = "province") %>% 
  # Create tooltip label 
  mutate(tooltip_label = paste0(province, ": ", total_cases))

china_map_plot <- ggplot(china_map_joined) +
  geom_sf(aes(fill = total_cases, text = tooltip_label)) + 
  theme_void() +
  labs(fill = "Number of Cases") + 
  theme(legend.position = "none")

china_map_plot_ggplotly <- 
  ggplotly(china_map_plot, tooltip = "text")

```

# Home

## Row {height=25%}
```{r}
value_box(
  title = "Total Cases",
  value = total_cases,
  showcase = bsicons::bs_icon("virus"),
  theme = value_box_theme(bg = "#518fd6"))
```

```{r}
value_box(
  title = "Total Hospitalisations",
  value = total_hospitalisation,
  showcase = bsicons::bs_icon("hospital"),
  theme = value_box_theme(bg = "#214773"))
```

```{r}
value_box(
  title = "Total Deaths",
  value = total_deaths,
  showcase = bsicons::bs_icon("file-earmark-medical"),
  theme = value_box_theme(bg = "#3f71ab"))
```

## Row {height=75%}

### {width=30%}
```{r tilte = "Number of Cases by Province"}
plot_flu
```

### {width=70%}
```{r "Map of Provinces by Number of Cases"}
china_map_plot_ggplotly

```

# Download Data

The data used in this dashboard is shown below and can be downloaded as a csv.

```{r}
htmltools::browsable(
  tagList(

    reactable(
      fluH7N9_china_2013,
      elementId = "fluH7N9_china_2013",
      searchable = TRUE, 
          filterable = TRUE),
    
    tags$button("Download as CSV", onclick = "Reactable.downloadDataCSV('fluH7N9_china_2013-table')"),
  )
)

```

# About

This dashboard analyzes data on the 2013 influenza A H7N9 outbreak in China.

Key points:

There were 136 total reported cases
62 cases were hospitalized
32 deaths occurred
The most affected provinces were Zhejiang, Shanghai and Jiangsu
The data is sourced from a Dryad dataset by Kucharski et al (2014) who collated case details from various reports. This dashboard provides an interactive visualization of the outbreak’s geographical distribution and key statistics.

Reference: Kucharski, A., Mills, H., Pinsent, A., Fraser, C., Kerkhove, M. V., Donnelly, C. A., & Riley, S. (2014). Distinguishing between reservoir exposure and human-to-human transmission for emerging pathogens using case onset data. PLOS Currents Outbreaks, 1. https://doi.org/10.1371/currents.outbreaks.e1473d9bfc99d080ca242139a06c455f

